Questions to be answered: 
\begin{itemize}
    \item Questions related to control strategies. (Divide this section in terms of different control strategies. Organize it.)
    \begin{enumerate}
    \item First of all what is control?
    \item What are the different control strategies. (signal/Loop-shaping/Numeric optimization)
    \item Why LQR for stabilization. What are its pros and cons.
    \item What is LPV? is it a control strategy or a modelling strategy?
    \item Why LPV? what are its pros and cons.
    \item How are the scheduling parameters identified for LPV? Search for some already present papers. Don't reinvent the wheel.
    \item What is MPC.
    \item Why MPC? How do we implement an MPC strategy. What are its requirements, pros and cons.
    \item what control strategy is used for swing up and why? What are the available options? Explain some literary work that is already available in this field.
    \end{enumerate}
    \item System modelling
    \begin{enumerate}
        \item What are the different ways in which a system is modelled and what is relevant to this thesis? 
        \item Pros and cons of first principle modelling 
        \item Pros and cons of data driven modelling 
        \item How are the parameters identified. List the parameters
        \item Lagrangian modelling
        \item Modelling from Data - Koopman
    \end{enumerate}
    \item Why is scaling important?
    \item Application of the above topics to ADIP.
    \begin{enumerate}
        \item 
    \end{enumerate}
\end{itemize}

First write about controllers. Swing-Up and Balancing control
Major questions:
\begin{itemize}
    \item How do we find the actuator bandwidth? How is this affected by the high frequency gain of controllers? What first of all is high frequency gain of controllers?
    \item How do we decide sampling rates? 
    \item What does it mean by -- Controller dynamics must be consistent with sampling rates? How does sampling rates affect bandwidth, dynamics etc. 
    
\end{itemize}
